---
title: "10-8-19 exploratory analysis"
author: "Megan Beaudry"
date: "10/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exploratory Analysis {.tabset}



##Prelimnary Loading
Load clean data
```{r}
clean_water_quality_data <- readRDS("~/Desktop/epid8060fall2019/MeganBeaudry-Project_WQ/data/processed_data/clean_water_quality_data.rds")
```

Load libraries
```{r}
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
```

##Exploratory Analysis - Summary Statistics
```{r}
summary(cleandata$Rainfall)
summary(cleandata$HF183)
summary(cleandata$HumM2)
summary(cleandata$CG01)
summary(cleandata$LeeSg)
summary(cleandata$Dog3)
summary(cleandata$MuBac)
summary(cleandata$Rum2Bac)
summary(cleandata$A_butlzeri_HSP60)
summary(cleandata$Salmonella_spp_InvA)
summary(cleandata$Campylobacter_spp_Van_Dkye)
summary(cleandata$Enterococcus_CCE)
summary(cleandata$total_coliforms)
summary(cleandata$E_coli)
summary(cleandata$Enteroalert)
summary(cleandata$thermotolerant_coliforms)
```

##Exploratory Analysis - Temporal Trends

*Temporal Trends All Pond Comparison*
Comparison with traditional water quality indicators and Arco
```{r}
ggplot(clean_water_quality_data) + 
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_bar(mapping = aes(x = Date_Sampled, y = Enterococcus_CCE), color = "purple", stat = "identity") + 
  ylab("CCE")
```

```{r}


ggplot(clean_water_quality_data) + 
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  ylab(expression("Log10_copies/100mL"))

ggplot(clean_water_quality_data) +
  geom_bar(mapping = aes(x = Date_Sampled, y = Enterococcus_CCE), color = "purple", stat = "identity") + 
  ylab("CCE")
  
```

The temporal trends by pond of HF183, HumM2 & A. butzleri to see if there was a relationship between source and pathogen
```{r}
All_Human_Arco <- ggplot(clean_water_quality_data) +
  geom_line(mapping = aes(x = Date_Sampled, y = HF183), color = "dark green") +
  geom_line(mapping = aes(x = Date_Sampled, y = HumM2), color = "purple") +
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_point(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_point(mapping = aes(x = Date_Sampled, y = HF183), color = "dark green") +
  geom_point(mapping = aes(x = Date_Sampled, y = HumM2), color = "purple") +
  labs (x = "Date Sampled", y = "Log10 Copies/100mL") +
  facet_wrap(~ Pond) +
  geom_hline(yintercept = 3.67, colour = "dark green", linetype = "dotted") +
  geom_hline(yintercept = 4.14, colour = "purple", linetype = "dotted") + 
  geom_hline(yintercept = 3.69, colour = "navy", linetype = "dotted") +
  theme_bw()
All_Human_Arco
```

```{r}
All_rain_Arco <- ggplot(clean_water_quality_data) +
  geom_line(mapping = aes(x = Date_Sampled, y = Rainfall), color = "dark green") +
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  labs (x = "Date Sampled", y = "Log10 Copies/100mL") +
  facet_wrap(~ Pond) +
  geom_hline(yintercept = 3.69, colour = "navy", linetype = "dotted") +
  theme_bw()
All_rain_Arco
  
```


*Temporal Trends grouped by Pond and Source*
Temporal trend of each pond individually with pathogen and source

Part 1: Human and Arco 
```{r}
McCall_Human_Arco <- clean_water_quality_data %>%
  filter(Pond == "McCall_Lake") %>%
  ggplot() +
  geom_line(mapping = aes(x = Date_Sampled, y = HF183), color = "dark green") +
  geom_line(mapping = aes(x = Date_Sampled, y = HumM2), color = "purple") +
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_hline(yintercept = 3.67, colour = "dark green", linetype = "dotted") +
  geom_hline(yintercept = 4.14, colour = "purple", linetype = "dotted") + 
  geom_hline(yintercept = 3.69, colour = "navy", linetype = "dotted") +
  geom_point(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_point(mapping = aes(x = Date_Sampled, y = HF183), color = "dark green") +
  geom_point(mapping = aes(x = Date_Sampled, y = HumM2), color = "purple") +
  labs (x = "Date Sampled", y = "Log10 Copies/100mL") +
  theme_bw()
McCall_Human_Arco
ggsave(McCall_Human_Arco, file = "McCall_Human_Arco.png")
```


All pathogens in McCall
```{r}
McCall_Pathogens <- clean_water_quality_data %>%
  filter(Pond == "McCall_Lake") %>%
  ggplot() +
  geom_line(mapping = aes(x = Date_Sampled, y = Campylobacter_spp_Van_Dkye), color = "dark green") +
  geom_line(mapping = aes(x = Date_Sampled, y = Salmonella_spp_InvA), color = "purple") +
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_hline(yintercept = 11.96, colour = "dark green", linetype = "dotted") +
  geom_hline(yintercept = 4.14, colour = "purple", linetype = "dotted") + 
  geom_hline(yintercept = 3.69, colour = "navy", linetype = "dotted") +
  geom_point(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_point(mapping = aes(x = Date_Sampled, y = Campylobacter_spp_Van_Dkye), color = "dark green") +
  geom_point(mapping = aes(x = Date_Sampled, y = Salmonella_spp_InvA), color = "purple") +
  labs (x = "Date Sampled", y = "Log10 Copies/100mL") +
  theme_bw()
McCall_Pathogens
ggsave(McCall_Pathogens, file = "McCall_Pathogens.png")
```

Took out Salmonella because we didnt find any...However THIS graph looks horrible becasue the LOD for campy is so high compared to the others (i.e., 11 to 3.5 on average). I need to come up with a better to show this if i want to show it. It really is a negative result, as we did not find much Campylobacter spp. However, they are highly related bacteria so some kind of relationship would be nice to show...
```{r}
McCall_Camp_Arco <- clean_water_quality_data %>%
  filter(Pond == "McCall_Lake") %>%
  ggplot() +
  geom_line(mapping = aes(x = Date_Sampled, y = Campylobacter_spp_Van_Dkye), color = "dark green") +
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_hline(yintercept = 11.96, colour = "dark green", linetype = "dotted") +
  geom_hline(yintercept = 3.69, colour = "navy", linetype = "dotted") +
  geom_point(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_point(mapping = aes(x = Date_Sampled, y = Campylobacter_spp_Van_Dkye), color = "dark green") +
  labs (x = "Date Sampled", y = "Log10 Copies/100mL") +
  theme_bw()
McCall_Camp_Arco
ggsave(McCall_Camp_Arco, file = "McCall_Camp_Arco.png")
```



```{r}
Country_Hills_Hum_Arco<- clean_water_quality_data %>%
  filter(Pond == "Country_Hills") %>%
  ggplot() +
  geom_line(mapping = aes(x = Date_Sampled, y = HF183), color = "dark green") +
  geom_line(mapping = aes(x = Date_Sampled, y = HumM2), color = "purple") +
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_hline(yintercept = 3.67, colour = "dark green", linetype = "dotted") +
  geom_hline(yintercept = 4.14, colour = "purple", linetype = "dotted") + 
  geom_hline(yintercept = 3.69, colour = "navy", linetype = "dotted") +
  geom_point(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_point(mapping = aes(x = Date_Sampled, y = HF183), color = "dark green") +
  geom_point(mapping = aes(x = Date_Sampled, y = HumM2), color = "purple") +
  labs (x = "Date Sampled", y = "Log10 Copies/100mL") +
  theme_bw()
ggsave(Country_Hills_Hum_Arco, file = "Country_Hills_Hum_Arco.png")
Country_Hills_Hum_Arco
```

```{r}
Inverness_Hum_Arco <- clean_water_quality_data %>%
  filter(Pond == "Inverness") %>%
  ggplot() +
  geom_line(mapping = aes(x = Date_Sampled, y = HF183), color = "dark green") +
  geom_line(mapping = aes(x = Date_Sampled, y = HumM2), color = "purple") +
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_hline(yintercept = 3.67, colour = "dark green", linetype = "dotted") +
  geom_hline(yintercept = 4.14, colour = "purple", linetype = "dotted") + 
  geom_hline(yintercept = 3.69, colour = "navy", linetype = "dotted") +
  geom_point(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_point(mapping = aes(x = Date_Sampled, y = HF183), color = "dark green") +
  geom_point(mapping = aes(x = Date_Sampled, y = HumM2), color = "purple") +
  labs (x = "Date Sampled", y = "Log10 Copies/100mL") +
  theme_bw()
Inverness_Hum_Arco
ggsave(Inverness_Hum_Arco, file = "Inverness_Hum_Arco.png")

```

Part 2: Birds and Arco 
```{r}
Inverness_Bird_Arco <- clean_water_quality_data %>%
  filter(Pond == "Inverness") %>%
  ggplot() +
  geom_line(mapping = aes(x = Date_Sampled, y = CG01), color = "dark green") +
  geom_line(mapping = aes(x = Date_Sampled, y = LeeSg), color = "purple") +
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_hline(yintercept = 3.4, colour = "dark green", linetype = "dotted") +
  geom_hline(yintercept = 3.58, colour = "purple", linetype = "dotted") + 
  geom_hline(yintercept = 3.69, colour = "navy", linetype = "dotted") +
  geom_point(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_point(mapping = aes(x = Date_Sampled, y = CG01), color = "dark green") +
  geom_point(mapping = aes(x = Date_Sampled, y = LeeSg), color = "purple") +
  labs (x = "Date Sampled", y = "Log10 Copies/100mL") +
  theme_bw()
Inverness_Bird_Arco
ggsave(Inverness_Bird_Arco, file = "Inverness_Bird_Arco.png")
```
```{r}
McCall_Bird_Arco <- clean_water_quality_data %>%
  filter(Pond == "McCall_Lake") %>%
  ggplot() +
  geom_line(mapping = aes(x = Date_Sampled, y = CG01), color = "dark green") +
  geom_line(mapping = aes(x = Date_Sampled, y = LeeSg), color = "purple") +
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_hline(yintercept = 3.4, colour = "dark green", linetype = "dotted") +
  geom_hline(yintercept = 3.58, colour = "purple", linetype = "dotted") + 
  geom_hline(yintercept = 3.69, colour = "navy", linetype = "dotted") +
  geom_point(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_point(mapping = aes(x = Date_Sampled, y = CG01), color = "dark green") +
  geom_point(mapping = aes(x = Date_Sampled, y = LeeSg), color = "purple") +
  labs (x = "Date Sampled", y = "Log10 Copies/100mL") +
  theme_bw()
McCall_Bird_Arco
ggsave(Inverness_Bird_Arco, file = "McCall_Bird_Arco.png")
```
```{r}
Country_Hills_Bird_Arco <- clean_water_quality_data %>%
  filter(Pond == "Country_Hills") %>%
  ggplot() +
  geom_line(mapping = aes(x = Date_Sampled, y = CG01), color = "dark green") +
  geom_line(mapping = aes(x = Date_Sampled, y = LeeSg), color = "purple") +
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_hline(yintercept = 3.4, colour = "dark green", linetype = "dotted") +
  geom_hline(yintercept = 3.58, colour = "purple", linetype = "dotted") + 
  geom_hline(yintercept = 3.69, colour = "navy", linetype = "dotted") +
  geom_point(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_point(mapping = aes(x = Date_Sampled, y = CG01), color = "dark green") +
  geom_point(mapping = aes(x = Date_Sampled, y = LeeSg), color = "purple") +
  labs (x = "Date Sampled", y = "Log10 Copies/100mL") +
  theme_bw()
Country_Hills_Bird_Arco
ggsave(Inverness_Bird_Arco, file = "Country_Hills_Bird_Arco.png")
```

*Temporal Trends One Marker Per Pond*
```{r}
McCall_HF183 <- clean_water_quality_data %>%
  filter(Pond == "McCall_Lake") %>%
  ggplot() +
  geom_line(mapping = aes(x = Date_Sampled, y = HF183), color = "dark green") +
  geom_point(mapping = aes(x = Date_Sampled, y = HF183), color = "dark green") + 
  labs (x = "Date Sampled", y = "Log10 Copies/100mL") +
  geom_hline(yintercept = 3.67, colour = "black", linetype = "dotted") +
  theme_bw() +
  scale_color_discrete(name = "Legend", labels = c("HF183", "LOD95"))
McCall_HF183
```

##Exploratory Analysis - Box and Whisker Plots
Boxplot of HF183 on all ponds
```{r}
All_Box_HF183 <- ggplot(data = clean_water_quality_data, aes(x = "", y = HF183)) + 
  geom_boxplot() +
  facet_wrap(~ Pond) +
  theme_bw() + 
  labs (x = "HF183", y = "Log10 Copies/100mL")
All_Box_HF183
```
Since McCall Lake has the most interesting data, lets look at the sampling sites within McCall Lake to see if one site has a higher median (is more contaminated)
```{r}
McCall_bysite_box_HF183 <- clean_water_quality_data %>%
  filter(Pond == "McCall_Lake") %>%
  ggplot(aes(x = "", y = HF183)) + 
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 6)) +
  facet_wrap(~ Sampling_Site) +
  theme_bw() +
  labs (x = "HF183", y = "Log10 Copies/100mL")
McCall_bysite_box_HF183
```

##Explatory Analysis - Violin Plots
The water quality data measured through molecular methods (i.e., Enterococcus CCE as determined by EPA Method 1611) would be best visualized using a violin plot
```{r}
McCall_Enterococcus_CCE_Violin <- clean_water_quality_data %>%
  filter(Pond == "McCall_Lake") %>%
  ggplot(aes(x = "", y = Enterococcus_CCE)) + 
  geom_violin() +
  geom_boxplot(width=0.1) +
  coord_cartesian(ylim = c(0, 6)) +
  facet_wrap(~ Sampling_Site) +
  theme_bw() +
  labs (x = "Enterococcus", y = "Log10 Copies/100mL")
McCall_Enterococcus_CCE_Violin
```

##Exploratory Analysis - Bar Charts
How many samples are positive for shigatoxin 1
```{r}
Shigatoxin1_bar <- clean_water_quality_data %>%
  ggplot() +
  geom_bar(aes(x = Shigatoxin_1), stat = "count", color = "magenta", fill = "white") +
  labs(x = "Shiga Toxin 1", y = "Counts") +
  theme(panel.background = element_rect(fill = "grey85"))
Shigatoxin1_bar
```
How many samples are positive for shigatoxin 2
```{r}
Shigatoxin2_bar <- clean_water_quality_data %>%
  ggplot() +
  geom_bar(aes(x = Shigatoxin_2), stat = "count", color = "yellow", fill = "white") +
  labs(x = "Shiga Toxin 1", y = "Counts") +
  theme(panel.background = element_rect(fill = "grey85"))
Shigatoxin2_bar
```

