---
title: "11-05-19 Univariate Anaysis"
author: "Megan Beaudry"
date: "11/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Univariate Analysis Script

Load the libraries to get the code to run. 

```{r load library}
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
library(dplyr)
library(cowplot)
library(grid)
```


Load the cleaned data that was produced from the cleaning script!
```{r load data, echo=FALSE}
clean_water_quality_data <- readRDS("../../data/processed_data/clean_water_quality_data.rds")
cleandata <- clean_water_quality_data 
```


Summary Statistics
Here are the summary statistics for each of the important variables. This will provide us with a good general overview of the data. 
```{r summary stats for all variables}
summary(clean_water_quality_data$Rainfall)
summary(clean_water_quality_data$HF183)
summary(clean_water_quality_data$HumM2)
summary(clean_water_quality_data$CG01)
summary(clean_water_quality_data$LeeSg)
summary(clean_water_quality_data$Dog3)
summary(clean_water_quality_data$MuBac)
summary(clean_water_quality_data$Rum2Bac)
summary(clean_water_quality_data$A_butlzeri_HSP60)
summary(clean_water_quality_data$Salmonella_spp_InvA)
summary(clean_water_quality_data$Campylobacter_spp_Van_Dkye)
summary(clean_water_quality_data$Enterococcus_CCE)
summary(clean_water_quality_data$total_coliforms)
summary(clean_water_quality_data$E_coli)
summary(clean_water_quality_data$Enteroalert)
```



Summary stats table for publication
```{r list of summary stats}
library(dplyr)
library(magrittr)
library(qwraps2)
options(qwraps2_markup = "markdown")

args(summary_table)

our_summary2 <-
  list("Human HF183" =
       list("min" = ~ min(cleandata$HF183),
            "max" = ~ max(cleandata$HF183),
            "median" = ~ median(cleandata$HF183)),
       "Human HumM2" =
       list("min" = ~ min(cleandata$HumM2),
            "median" = ~ median(cleandata$HumM2),
            "max" = ~ max(cleandata$Humm2)),
       "Arcobacter spp. HSP60" =
       list("min" = ~ min(cleandata$A_butzleri_HSP60),
            "max" = ~ max(cleandata$A_butzleri_HSP60),
            "median" = ~ median(cleandata$A_butzleri_HSP60)),
       "Canada Goose CG01" =
       list("min" = ~ min(cleandata$CG01),
            "max"  = ~ max(cleandata$CG01),
            "median" = ~ median(cleandata$CG01))
       )

whole <- summary_table(cleandata, our_summary2)
whole

#cannot get table to save
#png("../../results/summarystats.png", height=400, width=400)
#p<-tableGrob(whole)
#grid.arrange(p)
#dev.off()
```




BOX AND WHISKER PLOTS
We will make box and whisker plots to see the overall trends of all our variables. This will provide information on what may be contributing to A. butzleri. 

The that will be use Data: clean_water_quality_data

This first box and whisker plots we show us the mean of all the HF183 data facet wrapped by pond. This is helpful because it will show the varaiblity between the ponds. 
```{r Box and whisker HF183}
All_Box_HF183 <- ggplot(data = clean_water_quality_data, aes(x = "", y = HF183)) + 
  geom_boxplot() +
  facet_wrap(~ Pond) +
  theme_bw() + 
  labs (x = "HF183", y = "Log10 Copies/100mL")
All_Box_HF183
ggsave(filename = "../../results/All_Box_HF183.png", plot = All_Box_HF183)
```
After looking at the variablity between pond, I know now that I want to look at each pond by itself. McCall Lake looks very interesting which is unsurprising given what the other analysis has showed. 

```{r HF183 box and whisker by site then broken down by pond}
#labels are cut off in grid.arrange plot (title: HF183_boxplots_all_ponds)
#add legend saying what dotted line is 
library(gridExtra)
#make box and whisker for McCall
McCall_bysite_box_HF183 <- clean_water_quality_data %>%
  filter(Pond == "McCall_Lake") %>%
  ggplot(aes(x = "", y = HF183)) + 
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 6)) +
  facet_wrap(~ Sampling_Site) +
  theme_bw() +
  labs (x = "HF183", y = "Log10 Copies/100mL") +
 geom_hline(yintercept = 3.32, colour = "red", linetype = "dotted")

McCall_bysite_box_HF183
ggsave(filename = "../../results/McCall_bysite_box_HF183.png", plot = McCall_bysite_box_HF183)

#make box and whisker for Country HIlls
Country_Hills_bysite_box_HF183 <- clean_water_quality_data %>%
  filter(Pond == "Country_Hills") %>%
  ggplot(aes(x = "", y = HF183)) + 
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 6)) +
  facet_wrap(~ Sampling_Site) +
  theme_bw() +
  labs (x = "HF183", y = "Log10 Copies/100mL") +
  geom_hline(yintercept = 3.32, colour = "red", linetype = "dotted")
Country_Hills_bysite_box_HF183
ggsave(filename = "../../results/Country_Hills_bysite_box_HF183.png", plot = Country_Hills_bysite_box_HF183)

#make box and whisker for Inverness
Inverness_bysite_box_HF183 <- clean_water_quality_data %>%
  filter(Pond == "Inverness") %>%
  ggplot(aes(x = "", y = HF183)) + 
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 6)) +
  facet_wrap(~ Sampling_Site) +
  theme_bw() +
  labs (x = "HF183", y = "Log10 Copies/100mL") +
  geom_hline(yintercept = 3.32, colour = "red", linetype = "dotted")
Inverness_bysite_box_HF183
ggsave(filename = "../../results/Inverness_bysite_box_HF183.png", plot = Inverness_bysite_box_HF183)

#arrange them all together and save them as one png
HF183_boxplots_all_ponds <- grid.arrange(Inverness_bysite_box_HF183, Country_Hills_bysite_box_HF183, McCall_bysite_box_HF183)


ggsave(filename = "../../results/HF183_boxplots_all_ponds.png", plot = HF183_boxplots_all_ponds)

```

VIOLIN PLOTS
Violin plots will be used for the FIB (enterococcus and thermotolerant coliforms). I decided to use violin plots as they are commonly used in water quality data since the varation in the data can be so wide! It provides a good way to grasp what the dataset entails (i.e., do we see a lot of failures? Are they outliers? )

```{r Enterococcus_CCE_Violin}
#We need enterococcus to be italisized so it needs its own vector
italic_entero <- expression(paste(italic("Enterococcus")))

#violin plot McCall Lake
McCall_Enterococcus_CCE_Violin <- clean_water_quality_data %>%
  filter(Pond == "McCall_Lake") %>%
  ggplot(aes(x = "", y = Enterococcus_CCE)) + 
  geom_violin() +
  geom_boxplot(width=0.1) +
  coord_cartesian(ylim = c(0, 6)) +
  facet_wrap(~ Sampling_Site) +
  theme_bw() +
  labs (x = italic_entero, y = "Log10 CCE/100mL") +
  theme(axis.title.x = element_text( size = 7), axis.title.y = element_text( size = 7), legend.text = element_text( size = 4), legend.title=element_text(size=5)) +
  geom_hline(yintercept = 3.1, colour = "red", linetype = "dotted")

McCall_Enterococcus_CCE_Violin
ggsave(filename = "../../results/McCall_Enterococcus_CCE_Violin.png", plot = McCall_Enterococcus_CCE_Violin)

#violin plot Country Hills
Country_Hills_Enterococcus_CCE_Violin <- clean_water_quality_data %>%
  filter(Pond == "Country_Hills") %>%
  ggplot(aes(x = "", y = Enterococcus_CCE)) + 
  geom_violin() +
  geom_boxplot(width=0.1) +
  coord_cartesian(ylim = c(0, 6)) +
  facet_wrap(~ Sampling_Site) +
  theme_bw() +
  labs (x = italic_entero, y = "Log10 CCE/100mL") +
  theme(axis.title.x = element_text( size = 7), axis.title.y = element_text( size = 7), legend.text = element_text( size = 4), legend.title=element_text(size=5)) +
  geom_hline(yintercept = 3.1, colour = "red", linetype = "dotted")

Country_Hills_Enterococcus_CCE_Violin
ggsave(filename = "../../results/Country_Hills_Enterococcus_CCE_Violin.png", plot = Country_Hills_Enterococcus_CCE_Violin)

#violin plot Inverness
Inverness_Enterococcus_CCE_Violin <- clean_water_quality_data %>%
  filter(Pond == "Inverness") %>%
  ggplot(aes(x = "", y = Enterococcus_CCE)) + 
  geom_violin() +
  geom_boxplot(width=0.1) +
  coord_cartesian(ylim = c(0, 6)) +
  facet_wrap(~ Sampling_Site) +
  theme_bw() +
  labs (x = italic_entero, y = "Log10 CCE/100mL") +
  theme(axis.title.x = element_text( size = 7), axis.title.y = element_text( size = 7), legend.text = element_text( size = 4), legend.title=element_text(size=5)) +
  geom_hline(yintercept = 3.1, colour = "red", linetype = "dotted")

Inverness_Enterococcus_CCE_Violin
ggsave(filename = "../../results/Inverness_Enterococcus_CCE_Violin.png", plot = Inverness_Enterococcus_CCE_Violin)
#combine then all and save as one object
Entero_violin_all <- grid.arrange(Inverness_Enterococcus_CCE_Violin, Country_Hills_Enterococcus_CCE_Violin, McCall_Enterococcus_CCE_Violin)
```



```{r Thermotolerant Coliforms}
#violin plot McCall Lake
McCall_thermo_coliforms <- clean_water_quality_data %>%
  filter(Pond == "McCall_Lake") %>%
  ggplot(aes(x = "", y = thermotolerant_coliforms)) + 
  geom_jitter() +
  coord_cartesian(ylim = c(0, 6)) +
  facet_wrap(~ Sampling_Site) +
  theme_bw() +
  labs (x = "Total Coliforms", y = "CFU") +
  geom_hline(yintercept = 2.6, colour = "red", linetype = "dotted") +
  theme(axis.title.x = element_text( size = 7), axis.title.y = element_text( size = 7), legend.text = element_text( size = 4), legend.title=element_text(size=5)) 

McCall_thermo_coliforms
ggsave(filename = "../../results/McCall_thermo_coliforms .png", plot = McCall_thermo_coliforms)
#violin plot Country Hills
Country_Hills_thermo_coliforms <- clean_water_quality_data %>%
  filter(Pond == "Country_Hills") %>%
  ggplot(aes(x = "", y = thermotolerant_coliforms)) + 
  geom_jitter() +
  coord_cartesian(ylim = c(0, 6)) +
  facet_wrap(~ Sampling_Site) +
  theme_bw() +
  labs (x = "Total Coliforms", y = "CFU") +
  geom_hline(yintercept = 2.6, colour = "red", linetype = "dotted") +
  theme(axis.title.x = element_text( size = 7), axis.title.y = element_text( size = 7), legend.text = element_text( size = 4), legend.title=element_text(size=5)) 

Country_Hills_thermo_coliforms
ggsave(filename = "../../results/Country_Hills_thermo_coliforms.png", plot = Country_Hills_thermo_coliforms)
#violin plot Inverness
Inverness_thermo_coliforms <- clean_water_quality_data %>%
  filter(Pond == "Inverness") %>%
  ggplot(aes(x = "", y = thermotolerant_coliforms)) + 
  geom_jitter() +
  coord_cartesian(ylim = c(0, 6)) +
  facet_wrap(~ Sampling_Site) +
  theme_bw() +
  labs (x = "Total Coliforms", y = "CFU") +
  geom_hline(yintercept = 2.6, colour = "red", linetype = "dotted") +
  theme(axis.title.x = element_text( size = 7), axis.title.y = element_text( size = 7), legend.text = element_text( size = 4), legend.title=element_text(size=5)) 

Inverness_thermo_coliforms
ggsave(filename = "../../results/Inverness_thermo_coliforms.png", plot = Inverness_thermo_coliforms)

thermo_coliforms_all <- grid.arrange(Inverness_thermo_coliforms, Country_Hills_thermo_coliforms, McCall_thermo_coliforms)
```


COMBINING WATER QUALITY PLOTS
Lets use grid.arrange to combine our water qulaity by pond!
```{r Country Hills water quality}
McCall_WQ <- grid.arrange(Country_Hills_Enterococcus_CCE_Violin, Country_Hills_thermo_coliforms)
ggsave(filename = "../../results/Country_Hills_WQ_plots.png", plot = McCall_WQ)

```
```{r Inverness water quality}
McCall_WQ <- grid.arrange(Inverness_Enterococcus_CCE_Violin, Inverness_thermo_coliforms)
ggsave(filename = "../../results/Inverness_WQ_plots.png", plot = McCall_WQ)

```


```{r McCall water quality}
McCall_WQ <- grid.arrange(McCall_Enterococcus_CCE_Violin, McCall_thermo_coliforms)
ggsave(filename = "../../results/McCall_WQ_plots.png", plot = McCall_WQ)

```

BAR CHARTS

Shigatoxin producing E. coli is a categorical variable in this study, as it was identified with presence, absence, or NA when E. coli is not present so therefore STEC cannot be present (at this time that is what the literature states. It would not surprise me if STEC "jumped" to other bacteria at some point, so I am leaving it as such below). For this reason I wanted to to make bar plots to just get a general idea of the levels of shigatxoin. There are two genes for shigatoxin - STEC1 and STEC2. 

```{r STEC 1 bar chart}
Shigatoxin1_bar <- clean_water_quality_data %>%
  ggplot() +
  geom_bar(aes(x = Shigatoxin_1), stat = "count", color = "magenta", fill = "white") +
  labs(x = "Shiga Toxin 1", y = "Counts") +
  theme(panel.background = element_rect(fill = "grey85"))
Shigatoxin1_bar
ggsave(filename = "../../results/Shigatoxin1_bar.png", plot = Shigatoxin1_bar)
```

```{r STEC 2 bar chart}
Shigatoxin2_bar <- clean_water_quality_data %>%
  ggplot() +
  geom_bar(aes(x = Shigatoxin_2), stat = "count", color = "yellow", fill = "white") +
  labs(x = "Shiga Toxin 2", y = "Counts") +
  theme(panel.background = element_rect(fill = "grey85"))
Shigatoxin2_bar
ggsave(filename = "../../results/Shigatoxin2_bar.png", plot = Shigatoxin2_bar)
```

