---
title: "11-29-19 Exploratory Analysis"
author: "Megan Beaudry"
date: "12/2/2019"
output: html_document
---

---
title: "10-16-19 exploratory analysis"
author: "Megan Beaudry"
date: "10/8/2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exploratory Analysis {.tabset}



##Prelimnary Loading

Load clean data
```{r}
clean_water_quality_data <- readRDS("../../data/processed_data/clean_water_quality_data.rds")
```

Load libraries
```{r}
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
library(dplyr)
```

##Exploratory Analysis - Summary Statistics
```{r}
summary(clean_water_quality_data$Rainfall)
summary(clean_water_quality_data$HF183)
summary(clean_water_quality_data$HumM2)
summary(clean_water_quality_data$CG01)
summary(clean_water_quality_data$LeeSg)
summary(clean_water_quality_data$Dog3)
summary(clean_water_quality_data$MuBac)
summary(clean_water_quality_data$Rum2Bac)
summary(clean_water_quality_data$A_butlzeri_HSP60)
summary(clean_water_quality_data$Salmonella_spp_InvA)
summary(clean_water_quality_data$Campylobacter_spp_Van_Dkye)
summary(clean_water_quality_data$Enterococcus_CCE)
summary(clean_water_quality_data$total_coliforms)
summary(clean_water_quality_data$E_coli)
summary(clean_water_quality_data$Enteroalert)
summary(clean_water_quality_data$thermotolerant_coliforms)
```

##Exploratory Analysis - Temporal Trends

*Temporal Trends All Pond Comparison*
Comparison with traditional water quality indicators and Arco.
I am really unhappy with this figure. I need to brainstorm a better way to represent data
```{r}
ggplot(clean_water_quality_data) + 
  geom_line(mapping = aes(x = Date_Sampled, y = A_butzleri_HSP60), color = "navy") +
  geom_bar(mapping = aes(x = Date_Sampled, y = Enterococcus_CCE), color = "purple", stat = "identity") + 
  ylab("CCE")
```